---
title: "Lab 07 - Conveying the right message through visualisation"
author: "Insert your name here"
date: "Insert date here"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

### Exercise 1
```{r exercise-1-data}
library(tibble)
library(ggplot2)

df <- tribble(
  ~date, ~mask_count, ~nomask_count,
  "7/13/2020", 23, 9.5,
  "7/14/2020", 19.5, 9,
  "7/15/2020", 20, 9.5,
  "7/16/2020", 20.25, 9.9,
  "7/17/2020", 19.75, 9.75,
  "7/18/2020", 20, 9.5,
  "7/19/2020", 20.25, 9.25,
  "7/20/2020", 20.2, 8.75,
  "7/21/2020", 21, 8.5,
  "7/22/2020", 20.75, 8.5,
  "7/23/2020", 19.75, 8.5,
  "7/24/2020", 20.15, 9.15,
  "7/25/2020", 20, 9.9,
  "7/26/2020", 19.25, 10,
  "7/27/2020", 18.5, 9.85,
  "7/28/2020", 16.75, 9.75,
  "7/29/2020", 16.65, 9.8,
  "7/30/2020", 16.75, 9.9,
  "7/31/2020", 16.5, 9.6,
  "8/1/2020", 16.25, 9,
  "8/2/2020", 16, 9,
  "8/3/2020", 16, 9.15,
)

df_long <- df %>%
  pivot_longer(cols = c(mask_count, nomask_count), 
               names_to = "category", values_to = "count")

```
  

### Exercise 2

```{r exercise-2-plot}

ggplot(df_long, aes(x = as.Date(date, format="%m/%d/%Y"), y = count, 
                    color = category, group = category)) +
  geom_line(size = 1) +  
  geom_point(size = 2) +  
  labs(title = "Kansas COVID-19 7-day Rolling Average of Daily Cases",
       x = NULL,  # Correct way to remove x-axis label
       y = "Average Daily Cases (per 100,000 people)",
       color = "County Type") + 
  scale_color_manual(values = c("mask_count" = "blue", "nomask_count" = "red"),
                     labels = c("mask_count" = "Mask Mandate", 
                                "nomask_count" = "No Mask Mandate")) +  
  theme_minimal()
```
### Exercise 3
In my plot it is more clear that counties with mask mandates have higher daily averages compared with counties without mask mandates. In the original visualization the different scales for the y axis make it look as if there were lower daily averages in counties with mask mandates. The original visualization is misleading because it does not convey what the data shows.

### Exercise 4
This data simply shows that counties that have mask mandates have higher daily averages compared with counties that do not have mask mandates. This data does not show that mask mandates are ineffective because the daily averages could be the cause of other factors (e.g. population density)


### Exercise 5

The message conveyed by my plot is that counties with mask mandates have higher daily averages compared with counties without mask mandates. This is shown through using the same y axis for both lines. Also the legend is clearer compared to the first plot, making it easier to tell which line represents which category. Furthermore, the initial plot seems purposefully misleading with how it shifts the y axis scales to make it seem like counties with mask mandates have lower daily averages.

### Exercise 6
For a plot to convey the opposite message that would mean that it would seem that the mask mandate counties have lower daily averages compared with counties without mask mandates.

To do this, i could instead focus on the changing rates rather than the averages because the mask mandate counties have a higher rate of change compared with the counties without mask mandates. This would make it seem like the mask mandate counties have lower daily averages compared with counties without mask mandates.

### Exercise 7

```{r exercise-7-mutate}
library(dplyr)

df_long <- df_long %>%
  arrange(category, as.Date(date, format="%m/%d/%Y")) %>%  
  group_by(category) %>%  
  mutate(first_day_count = count[date == "7/13/2020"],  
         change_from_first = count - first_day_count) %>%  
  ungroup()
```


```{r exercise-7-plot}
ggplot(df_long, aes(x = as.Date(date, format="%m/%d/%Y"), y = change_from_first, 
                    color = category, group = category)) +
  geom_line(size = 1) +  
  geom_point(size = 2) +  
  labs(title = "Kansas COVID-19 7-day Rolling Average of Daily Cases",
       x = NULL,  # Correct way to remove x-axis label
       y = "Change in Daily Cases (per 100,000 people)",
       color = "County Type") +  
  scale_color_manual(values = c("mask_count" = "blue", "nomask_count" = "red"),
                     labels = c("mask_count" = "Mask Mandate", 
                                "nomask_count" = "No Mask Mandate")) +  
  theme_minimal()
```

This plot has change in daily cases on the y axis and date on the x axis. The change is the difference from the first day. Because the mask mandate counties have a higher rate of change compared with the counties without mask mandates, it makes it seem like the mask mandate counties have lower daily averages compared with counties without mask mandates. This plot conveys the opposite message compared to the accurate plot.